### 2020-04-22 README.Rmd
vignette

A dataset that includes COVID-19 data and **all the other demographic variables** can be obtained executing `getus_all` (get-U.S-all) and `getit_all` (get-Italy-all)

@@

# to retrive u.s data
all_us <- getus_all()

# to retrive italy data
all_it <- getit_all()
```

@@

The function `get  all` executes

`get  covid` and joins the resulting `tibble` with datasets from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by PlaceIQ.

In particular, this is a list of datasets included:

* `act_ind`: county-level device exposure index (DEX), an index of activity.
* `acm_househ`: several metric regarding household composition (2018).
* `age_sex`: age and sex distribution (2018).
* `fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
* `hospbeds`: total hospital beds in each county (2019).
* `mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of

    * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
    * several type of cancer.
    * emergency,  medical admissions and annual visits.
    * pneumoccocal vaccine.
    * tabacco use.

*Note that* info for some counties are missing in some datasets. For example, `hospbeds` contains info on 2545 counties, `fl65` has 3224 counties, whereas datasets from the Census have 3220 counties.

@@

The function `get_all` executes `get_covid` and aggregates the resulting `tibble` with some of the datasets below and, normalizes some of the variables:

* `act_it`:  Change in  retail and recreation activity as reported by google [not joined with `get_all()`  but accessible].
* `bweight_it`: Body mass index in regions of Italy, in the general population.
* `cancer_it`: Number of cancer patients in each region by type.
* `chronic_it`: Number of people suffering of chronic conditions by region and type.
* `dem_it`: Number of people by region, sex, age.
* `firstaid_it`: Number of people using first aid or medical guard (3 months). 
* `fl_it`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
* `fl_it_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
* `hospbed_it`: Inpatient hospital beds per 1000 people.
* `house_it`:  household crowding index from 2014 to 2018 in each region.
* `netinc_it`:  median net annual households income (including imputed rents, in euros).
* `pm2.5_it`:  emission of particulate 2.5 in tons per region.
* `regions_area`: Area in square meters of each region. Used to calculate density per region.
* `smoking_it`: Number of people aged 14 years and over that self-refered as smoker, non smoker, or past smoker by region.

## Raw Data and other scripts

>>>>

A dataset that includes COVID-19 data and **all the other demographic variables** can be obtained executing `getus_all` (get-U.S-all) to get U.S data, or `getit_all` (get-Italy-all) to get Italy data.

@@

# to retrieve u.s data
all_us <- getus_all()

# to retrieve italy data
all_it <- getit_all()
```

Details on the other functions and the datasets can be found in the Vignette of the package.

@@

Data regarding COVID-19 number of cases and deaths (at the county level) were obtained from the [New York Time git repository](https://github.com/nytimes/covid-19-data), whereas number of tests and hospitalizations from the [the Covid Tracking Project].

The other datsets were obtained from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by `PlaceIQ`.

This is a list of U.S. datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:

* `act_ind`: county-level device exposure index (DEX), an index of activity.
* `acm_househ`: several metric regarding household composition (2018).
* `age_sex`: age and sex distribution (2018).
* `fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
* `hospbeds`: total hospital beds in each county (2019).
* `mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
    * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
    * several type of cancer.
    * emergency,  medical admissions and annual visits.
    * pneumoccocal vaccine.
    * tobacco use.

*Note that* info for some counties are missing in some datasets. For example, `hospbeds` contains info on 2545 counties, `fl65` has 3224 counties, whereas datasets from the Census have 3220.

@@

Data regarding COVID-19 deaths, confirmed cases, test and hospitalizations are extracted from the [Protezione Civile repository](https://github.com/pcm-dpc/COVID-19)

Demographic metrics and other stats regarding medical conditions and vaccinations were obtained from the [ISTAT](http://dati.istat.it/?lang=en), [Ministero della Salute](http://www.salute.gov.it/portale/home.html). Data regarding `pm2.5` were collected by [Istituto Superiore Per La protexione Ambientale](https://annuario.isprambiente.it/pon/basic/14)

This is a list of Italy datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:

* `bweight_it`: Body mass index in regions of Italy, in the general population.
* `cancer_it`: Number of cancer patients in each region by type.
* `chronic_it`: Number of people suffering of chronic conditions by region and type.
* `dem_it`: Number of people by region, sex, age.
* `firstaid_it`: Number of people using first aid or medical guard (3 months). 
* `fl_it`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
* `fl_it_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
* `hospbed_it`: Inpatient hospital beds per 1000 people.
* `house_it`:  household crowding index from 2014 to 2018 in each region.
* `netinc_it`:  median net annual households income (including imputed rents, in euros).
* `pm2.5_it`:  emission of particulate 2.5 in tons per region.
* `regions_area`: Area in square meters of each region. Used to calculate density per region.
* `smoking_it`: Number of people aged 14 years and over that self-referred as smoker, non smoker, or past smoker by region.

## Raw Data and other Scripts
